---
title: "Prediction of Backorders"
author: "Jake Close"
date: "11/16/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Problem Context:

Backordering boils down to having orders that a company cannot fulfill or more orders than you have on stock. For most businesses, having this many sales is fantastic, but can be a huge problem if its not handled properly. If a company is consistently running on backorder, it could a sign of poor inventory management and the company could be losing out on business by not providing the products customers want. If a company can identify products or parts at risk of going to back order before the event occurs, they can properly react.



##Data:
Data is provided from Kaggle.com (https://www.kaggle.com/tiredgeek/predict-bo-trial)
Training data file contains the historical data for the 8 weeks prior to the week we are trying to predict. The data was taken as weekly snapshots at the start of each week.


###Importing Data and Libraries
```{r}
library(dummies)
library(class)
library(gmodels)
library(ggplot2)
library(dplyr)
library(C50)
orders_test<- read.csv("Kaggle_Test_Dataset_v2.csv")
orders_train <- read.csv("Kaggle_Training_Dataset_v2.csv")
```



##Data Cleaning

Lets analyze the structure of our data. We find there are 1,687,861 observations and 23 variables. Our predictor variable is "went_on_backorder". At first glance, there are are mixture of factors, integer and numerical columns. Also, some variables have NA's which we must take into account, and many variables appear to be sparse.
```{r}
str(orders_train)
```

```{r}
summary(orders_train)
```
There is an issue when reading in the last row, lets remove it. Lets also clean the data of NAs


```{r}
orders_train = orders_train[1687861,]

colSums(is.na(orders_train))

```

Looks like we have some NA values for lead time. And a singular row of NA for some other factors. Lets set the lead time factor to the average and remove the rows with the NAs for the other factors.
```{r}
orders_train$lead_time[is.na(orders_train$lead_time)] <- median(orders_train$lead_time,na.rm=T)
orders_train <- orders_train[!row.has.na,]
orders_train <- na.omit(orders_train)
colSums(is.na(orders_train))
```

Much better. Next, lets observe the distribution of our predictor variable:
```{r}

round(prop.table(table(orders_train$went_on_backorder)) * 100, 1)

```
We see that the majority of the products do not go on backorder. This means that we must have a very fined tune algorithm to to detect these small variances.

Lets use a decision tree to train a model.
```{r}
backorder_model <- C5.0(orders_train[,-23], orders_train$went_on_backorder)

```

